---
title: "DV 1-3 Murray Homework"
date: October 1, 2023
author: "Ethan Murray"
format: 
    html: 
      self-contained: true
      theme: spacelab
      embed-resources: true
      toc: true
      number-sections: false
      code-tools: true
      code-fold: false
      df-print: kable
editor: source
execute:
  echo: true
  message: true
  warning: true
editor_options: 
  chunk_output_type: inline
---
```{r}
#| warning: false
library(tidyverse)
```
# Question 1 
Before looking at the data, I would expect that people who record more exercise will have a lower BMI and vice versa.

```{r}
#| warning: false
exercise_data <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/visualize_data.csv")
glimpse(exercise_data)
```

```{r}
cor(exercise_data$Exercise, exercise_data$BMI)
```

A correlation value of -.06447185 suggests a negative weak correlation betweeen BMI and Exercise score. 

```{r}
ggplot(data = exercise_data, mapping = aes(x = Exercise, y = BMI)) + 
     geom_point()
```

Definitely not the expected scatterplot... 

# Question 2 

```{r}
#| warning: false
library(causact)
glimpse(corruptDF) 
```

```{r}
?corruptDF
```

The CPI2017 represents a numerical score of the level of corruption within a country (0 = highly corrupt), and the HDI2017 represents a numerical score of the countries level of development including education and income. 

# Question 3 

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017)) + 
          geom_point()
```
Observing the scatter plot above we can see a positive seemingly linear relationship between HDI and CPI. This suggests that as the human development index score increases the corruption score increases which identifies a lower level of corruption. 

# Question 4 

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017)) + 
                 geom_point() + geom_smooth(method = "lm")
```

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017)) + 
                 geom_point() + geom_smooth(method = "gam")
```
After using both the "lm" and "gam" smoothing methods, I prefer the "gam" method. Comparing the scatter plot before using geom_smooth, the "gam" method seemed to capture the slightly exponential relationship bewteen HDI and CPI better than the "lm" method did. 

# Question 5 

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017)) + 
                       geom_point(mapping = aes(color = region)) + geom_smooth(method = "gam")
```
Although some patterns begin to emerge when introducing a color aesthetic for  region, the patterns are not jumping off the graph at first glance. 

The next step to understand the patterns, is to introduce a facet wrap function by region. The code below produced a scatterplot for each of the regions. Displaying trends and differences in a clear easier to understand manner. 

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017)) + 
          geom_point(mapping = aes(color = region)) + facet_wrap("region")
```
The SSA region clearly displays the lowest HDI score as well as the lowed CPI suggesting low development and high amounts of corruption. On the contrary, EU W. Europe displays a condensed, high HDI scores and higher CPI scores suggesting high levels of development and low levels of corruption. 

# Question 6 

```{r}
ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017)) + 
                geom_point(mapping = aes(color = region)) + facet_wrap("region") + scale_x_reverse()
```
Before reversing the X-Axis a "good" country with low corruption and high development would appear in the top right of the graph. After reversing the x-axis, a country with low corruption and high development would appear in the top left of the graph. 

# Question 7 

```{r}
 ggplot(data = corruptDF, mapping = aes(x = HDI2017, y = CPI2017)) +
    geom_point(mapping = aes(color = region)) +
          facet_wrap("region") +
          labs(x = "Human Development Index", y = "Corruption Perceptions Index",
                          title = "HDI and CPI",
                          subtitle = "Data points are Countries",
                         caption = "Source: Transparency International.")
```
# Question 8

```{r}
ggsave(filename = "HDI_CPI.png")
```