---
title: "DV5"
date: October 5, 2023
author: "Ethan Murray"
format: 
    html: 
      self-contained: true
      theme: spacelab
      toc: true
      number-sections: false
      code-tools: true
      code-fold: false
      df-print: kable
editor: source
execute:
  echo: true
  message: true
  warning: true
editor_options: 
  chunk_output_type: inline
---

```{r}
#| message: false
library(tidyverse)
```

```{r}
#| message: false
wncaa <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/wncaa.csv")
```

```{r}
glimpse(wncaa)
```

# Question 1

```{r}
champ_ncaa <- wncaa %>% 
  filter(tourney_finish == "Champ")
```

```{r}
by_school_champ <- champ_ncaa %>% 
  group_by(school) %>% 
  summarize(tourney_finish = n())
```

or from the book:

```{r}
champ_by_school <- champ_ncaa %>%
    group_by(school) %>%
    summarize(tourney_finish = n()) %>%
    mutate(freq = tourney_finish / sum(tourney_finish),
           pct = round((freq*100), 0))
```

```{r}
Bar_Plot_Schools <- ggplot(champ_by_school, aes(x = school, y = pct)) +
geom_col(position = "dodge2") +
    labs(x = "School",y = "Percent") +
    theme(legend.position = "top") +
  coord_flip()
Bar_Plot_Schools
```

Some patterns become pretty evident from the bar plot... Uconn and Tennesee have the lions share of championships with a combined percentage of 52%, with no other school above 5% of championships.

# Question 2

```{r}
winners <- wncaa %>% 
     filter(school %in% champ_ncaa)
```

```{r}
ggplot(data = winners,
            aes(x = school, y = seed)) +
geom_boxplot() + coord_flip()
```

```{r}
boxplot_schools <- ggplot(data = winners, mapping = aes(x = reorder (school, seed), y = seed)) +
geom_boxplot() + coord_flip() + geom_jitter(position = position_jitter(width = .15))
boxplot_schools
```

```{r}
by_school <- winners %>% group_by(school) %>%
    summarize(seed_mean= mean(seed, na.rm = TRUE),
              seed_sd = sd(seed, na.rm = TRUE))
```

```{r}
pointplot_schools <- ggplot(data = by_school, mapping = aes(x = school, y = seed_mean)) +
  geom_point(size = 3) + 
  labs(x = "School",
       y = "Average Seed") +
  theme(legend.position = "top") + coord_flip()
pointplot_schools
```

I was honestly surprised to see how Tennessee had a higher average seed than Uconn despite Uconn having a higher share of championships in the same time period.

```{r}
violin_schools <- ggplot(data = winners, mapping = aes(x = reorder (school, seed), y = seed)) +
geom_violin() + coord_flip() + geom_jitter(position = position_jitter(width = .15))
violin_schools
```

I think the boxplot merged with a average mean summary of the data is more useful in understanding both the overall range/ spread of the data but also where the highest concentrations of seeds are relative to the rest of the teams.

# Question 3

```{r}
other_pointplot <- ggplot(data = winners, mapping = aes(x = reorder (school, seed), y = seed)) +
geom_point() + coord_flip() + geom_jitter(position = position_jitter(width = .15))
other_pointplot
```

I'm not a fan of this plot personally, but when I used geompoint to look at the averages of the teams previosly I found it to be more useful. It's hard to understand the spread of the data and where the respective quartiles are concentrated.

# Question 4

```{r}
winners_mean_sd <- winners |> 
  group_by(school) |> 
  summarize(across(where(is.numeric),
                   list(mean = mean,
                        sd = sd)))
```

```{r}
reordered_pointplot <- ggplot(data = winners_mean_sd, mapping = aes(x = reorder (school, reg_percent_mean), y = reg_percent_mean)) +
geom_point() + coord_flip() + geom_jitter(position = position_jitter(width = .15))
reordered_pointplot
```

Texas A&M has the lowest average regular season win percentage for any of the champions since 1982. The lowest reg win percentage of a championship team was Tennessee in 1997 when they won 69.7% of reg games.

```{r}
pointrange_plot <- ggplot(data = winners_mean_sd, mapping = aes(x = reorder(school, reg_percent_mean), y = reg_percent_mean)) + geom_pointrange(mapping = aes(ymin = reg_percent_mean - reg_percent_sd, ymax = reg_percent_mean + reg_percent_sd)) + coord_flip()
pointrange_plot
```

The school with the narrowest interval is Texas A&M. This means they consistently win roughly 67%-79% of regular season games without much variance season by season.

```{r}
nomeanplot <- ggplot(data = winners_mean_sd, mapping = aes(x = reorder(school, reg_percent_mean), y = reg_percent_mean)) + geom_linerange(mapping = aes(ymin = reg_percent_mean - reg_percent_sd, ymax = reg_percent_mean + reg_percent_sd)) + coord_flip()
nomeanplot
```

I don't think this is quite as good as the pointrange, because while it does display a range of percentages, the point for the mean is great for the visual perception of where the mean lies.
